Reaction-diffusion spacial modeling of COVID-19 in Chicago

### 2021-10-26 draft/chicago-spatial-covid-draft.tex
Added PDE section
Fixed deltas to nablas
Lots of updates.

The population model at the PDE level is an autonomous diffusion with a source:
\begin{align}
	S_t	&=	\Delta (\D_s \Delta S) - \beta SI,	\\
	I_t	&=	\Delta (\D_I \Delta I) + \beta SI - \gamma I,	\\
	R_t	&=	\gamma I.
\end{align}
In principle the $R$ population can have a diffusion, but since we assume this population has immunity we assign $\D_R = 0$.

@@

The next step is to identify the parameters at the ODE level.
To do this, we use a nonlinear optimization algorithm.
We determine the optimal parameters by minimizing the Euclidean distance between the time series generated by the model ($\num$) and the corresponding observed ($\obs$) time series,
\begin{equation} \label{eq:obj}
	\N = \sum_i^{t_\text{fit}^\text{end}} \left( \left| \log(C_\num (t_i)) - \log(C_\obs (t_i)) \right|^2 + \left| \log(D_\num (t_i)) - \log(D_\obs (t_i)) \right|^2 \right)
\end{equation}
where the index $i$ identifies a point in the time series.
The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of deceased $D(t) = R(t)$.

@@

\appendix
\section{ODE Dynamics}
Here we briefly discuss some dynamics of the ODE system.
Note that for the simply SIR model as we have here, $R$ is completely determined when $S$ and $I$ are known, so we need only consider
\begin{align*}
S_t	&=	- \tilde{\beta} SI, \\
I_t	&=	\tilde{\beta} SI - \gamma I,
\end{align*}
where $\tilde{\beta} = \beta / N$.

>>>>

The population model at the PDE level is an autonomous diffusion with a source:
\begin{align}
	S_t	&=	\nabla (\D_s \nabla S) - \beta SI,	\\
	I_t	&=	\nabla (\D_I \nabla I) + \beta SI - \gamma I,	\\
	R_t	&=	\gamma I.
\end{align}
In principle the $R$ population can have a diffusion, but since we assume this population has immunity we assign $\D_R = 0$.

@@

The next step is to identify the parameters at the ODE level.
To do this, we use a nonlinear optimization algorithm.
We determine the optimal parameters by minimizing the Euclidean distance between the time series generated by the model ($\num$) and the corresponding observed ($\obs$) time series,
\begin{equation} \label{eq:obj}
	\N = \sum_i^{t_\text{fit}^{\text{end}}} \sqrt{ (C_\num (t_i) - C_\obs (t_i))^2 + (D_\num (t_i) - D_\obs (t_i))^2}
\end{equation}
where the index $i$ identifies a point in the time series.
The parameters are optimized to reproduce the reported total number of infected cases $C(t) = I(t) + R(t)$, and the total number of deceased $D(t) = R(t)$.

@@

\subsection{PDE Model}
We write the system in the form of a reaction-diffusion equation
\begin{equation} \label{eq:rd}
	\partial_t \mathbf{u} = \D \nabla^2 \mathbf{u} + \mathbf{R (u)}
\end{equation}
where $\mathbf{u} = \mathbf{u}(x,y,t)$ is a vector of density variables,  $\mathbf{R(u)}$ is a nonlinear term describing local reaction kinetics, and $\D$ denotes a diagonal diffusion coefficient matrix.
In this case we have
\begin{align*}
	\mathbf{u} &= \begin{pmatrix} S(x,y,t) \\ I(x,y,t) \\ R(x,y,t) \end{pmatrix} &
	\mathbf{R(u)} &= \begin{pmatrix} - \beta S I \\ \beta S I - \gamma I \\ \gamma I \end{pmatrix} &
	\D &= \begin{pmatrix} \D_S & 0 & 0 \\ 0 & \D_I & 0 \\ 0 & 0 & 0 \end{pmatrix}
\end{align*}
\appendix
\section{ODE Dynamics}
Here we briefly discuss some dynamics of the ODE system.
Note that for the simply SIR model as we have here, $R$ is completely determined when $S$ and $I$ are known, so we need only consider
\begin{align*}
S_t	&=	- \tilde{\beta} SI, \\
I_t	&=	\tilde{\beta} SI - \gamma I,
\end{align*}
where $\tilde{\beta} = \beta / N$.

### 2021-10-26 pres-8-27/slides/social-01.tex
SoReMo updates
Lots of updates.

\begin{slide}{Social and Community Connections}
	{\large Original context:} Working with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\

@@

	What are the effects of \textbf{age} and \textbf{race/ethnicity} in the Chicago area?
\end{slide}

>>>>

\begin{slide}{Social and Community Connections}
	{\large Summer 2020 context:} IPRO with David Eads (NPR) to identify \textit{demographic trends} in COVID-19. \\

@@

	\begin{itemize}
		\item How does the impact of COVID-19 (in Chicago) depend on \textbf{age} and \textbf{race/ethnicity}?
	\end{itemize}

	\vspace{1cm}

	{\large Present context:} COVID-19 is still here. Where is it impacting and why?

	\begin{itemize}
		\item Does \textbf{community structure} play a role?
		\item How effective are \textbf{vaccines}? (Or what is thwarting them?)
		\item Are \textbf{care facilities} properly allocated?
		\item \emph{Other ideas?}
	\end{itemize}

	\vspace{.7cm}

	{\large \textbf{Goal:} Can we design a model to answer these questions?}
\end{slide}
